---
title: "Massachusetts Crimes Master List"
output: rmarkdown::github_document
---

[Citizens for Juvenile Justice](https://www.cfjj.org) (CFJJ) have partnered [Code for Boston](https://www.codeforboston.org) (CfB) to develop a tool to streamline the expungement process for folks who have a criminal record before the age of 21 in Massachusetts.  

This project is exploratory data analysis on the 
We have a file of all crimes from the State of Massachuestts that we need to do an initial analysis on for the Clean Slate project sponsored by [Citizens for Juvenile Justice](https://www.cfjj.org). We will be running summary statistics on it to see the break out of the categories versus FBI categories. Are there ways to neatly, easily, remove entire categories of crimes from further analysis?  

packages for analysis
```{r echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(DataExplorer)
library(SmartEDA)


```

# Data Source  
The data for this project was provided by MA Youth Services and coded at 
```{r include=FALSE}
ma_crimes <- 
  read.csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vRQtMFa4KZlBlXpklfT9uSE383NaRtN8C5-L1FeSS1sTxX0Ws3Xam6Ahbpuf2BSOzR1Ugc9AZGPRLlu/pub?gid=579055210&single=true&output=csv", stringsAsFactors=FALSE) %>% 
  clean_names() %>% 
  na_if("#N/A") %>% 
  mutate(expungeable = replace(expungeable, expungeable != "Yes", "No"),
         fbi_category = replace(fbi_category, fbi_category == c("synthetic Narcotics – Manufactured Narcotics which can cause true drug addiction (Demerol, Methadones)"), c("Synthetic Narcotics – Manufactured Narcotics which can cause true drug addiction (Demerol, Methadones)")),
         fbi_category = replace(fbi_category, fbi_category == c("forgery and Counterfeiting"), c("Forgery and Counterfeiting")))

```

So thus far, we've read in the data, converted the headers to snake case, and replaced the `#N/A`s with proper `NA`. I may return to this decision and find another way to replace those values.

But we're probably going to need to change more things. Perhaps consider:

 - replace "#N/A" with `NA`  
 - adust binary fields to either `TRUE`/`FALSE` or logit `1`/`0`.
    - `expungible`  <= _as an aside, the variations in syntax are forcing different values. D:_
    - `police_notification`
    - `sorb`
    - `sdp`
    - `now_legal`  
 - should all these variables remain as factors? idk, seems like not a good idea.  
 
 With that said, I have some concerns / wishlists.  
 
  - proper, tidy data:  
    - section and chapter columns   
    - uniform descriptions   
    - confirming the categories that were used. Heard they weren't the correct ones? 
 - double check indent

```{r message=TRUE, warning=TRUE, include=FALSE, paged.print=TRUE}
#DataExplorer
create_report(ma_crimes)

```

preview
```{r}

summary(ma_crimes)
glimpse(ma_crimes) 

```


#univariate analysis 

```{r}

```



two-way table
```{r}
fbi_exp <- ma_crimes %>% 
  tabyl(fbi_category, expungeable) %>% 
  adorn_totals(c("row", "col"))

str(fbi_exp)

```


visualize `fbi_exp`
```{r}

```

Felony vs Misdemeanor
```{r}
ma_crimes %>% group_by(penalty_type) %>% summarise(n())

fbi_exp %>% adorn_ns()


```

```{r}

ExpData(data = ma_crimes, type = 2)

```

